---
title: ''
output:
  pdf_document: default
  html_document: default
---

```{r}
library(dplyr)
```

```{r}
df <- read.csv("ny.csv")
```

```{r}
plot(sort(df$WAGP))
```

```{r}
hist(df$WAGP)
```


```{r}
table(df$NOC)
```
Drop rows with NOC = -1 

```{r}
counts_wkhp <- table(df$WKHP)
barplot(counts_wkhp,xlab="Hours Worked per Week")
```

Create categories for part time [0, 35) full time [35-50] and overtime (50+)

```{r}
sum(df$WKHP < 35) / nrow(df) * 100
sum(df$WKHP >= 35 & df$WKHP <= 50) / nrow(df) * 100
sum(df$WKHP > 50) / nrow(df) * 100
```

```{r}
# "b" housing tenure - N/A
table(df$TEN)
```

Drop rows with TEN = b

```{r}
# "b" vehicles - N/A
table(df$VEH)
```
Drop rows with VEH = b

```{r}
# HH presence and age of children "b" N/A, "4" no children "1" only < 6, "2" only 6-17, "3" both 1 and 2
table(df$HUPAOC)
```

drop rows with HUPAOC = b

```{r}
table(df$WIF)
sum(df$WIF == "b") / nrow(df) * 100
```

Missing value is a considerable portion of data -> drop variable for now
in the future, possible to split into binary category (has_WIF) with values

```{r}
# drop since COW was already filtered
# table(df$COW)
```

```{r}
# remap "b" to no
table(df$FER)
```

```{r}
df %>%
  mutate(Absent = ifelse(ESR == 2, 1, 0)) %>%
  mutate(NonAbsent = ifelse(ESR == 1, 1, 0)) %>%
  mutate(Absent.Weighted = ifelse(ESR == 2, 1, 0) * PWGTP) %>%
  mutate(NonAbsent.Weighted = ifelse(ESR == 1, 1, 0) * PWGTP) %>%
  mutate(Absent.Weighted.log = ifelse(ESR == 2, 1, 0) * ceiling(log(PWGTP))) %>%
  group_by(NAICS) %>%
  summarise(
    Percent.Unweighted = sum(Absent) / (sum(Absent) + sum(NonAbsent)) * 100,
    Percent.Weighted = sum(Absent.Weighted) / (sum(Absent.Weighted) + sum(NonAbsent.Weighted)) * 100,
    Percent.Oversample = sum(Absent.Weighted) / (sum(Absent.Weighted) + sum(NonAbsent)) * 100,
    Percent.Oversample.log = sum(Absent.Weighted.log) / (sum(Absent.Weighted.log) + sum(NonAbsent)) * 100
    )
```

```{r}
df_clean <- df %>%
  select(-grep("PWGTP\\d", colnames(df))) %>%  # drop extra weight columns
  select(-ends_with("_label")) %>%  # drop labels
  select(-c("WGTP", "OCCP", "WKWN", "JWTRNS", "ST", "RT", "NAICSP")) %>%
  filter(NOC != -1) %>%
  mutate(WKHP.PartTime = ifelse(WKHP < 35, 1, 0)) %>%
  mutate(WKHP.FullTime = ifelse(WKHP >= 35 & WKHP <= 50, 1, 0)) %>%
  mutate(WKHP.OverTime = ifelse(WKHP > 50, 1, 0)) %>%
  filter(TEN != -1) %>%
  mutate(TEN.OwnLoan = ifelse(TEN == 1, 1, 0)) %>%
  mutate(TEN.Own = ifelse(TEN == 2, 1, 0)) %>%
  mutate(TEN.Rent = ifelse(TEN == 3, 1, 0)) %>%
  mutate(TEN.NoPay = ifelse(TEN == 4, 1, 0)) %>%
  filter(VEH != "b") %>%
  mutate_at("VEH", as.integer) %>%
  filter(HUPAOC != "b") %>%
  mutate(HUPAOC.UnderSix = ifelse(HUPAOC == 1, 1, 0)) %>%
  mutate(HUPAOC.OverSix = ifelse(HUPAOC == 2, 1, 0)) %>%
  mutate(HUPAOC.Both = ifelse(HUPAOC == 3, 1, 0)) %>%
  mutate(HUPAOC.None = ifelse(HUPAOC == 4, 1, 0)) %>%
  mutate(FER = ifelse(FER == 1, 1, 0)) %>%
  mutate(HINS1 = ifelse(HINS1 == 1, 1, 0)) %>%
  mutate(MALE = ifelse(SEX == 1, 1, 0)) %>%
  mutate(DIS = ifelse(DIS == 1, 1, 0)) %>%
  mutate(Absent = ifelse(ESR == 2, 1, 0)) %>%
  mutate(HICOV = ifelse(HICOV == 1, 1, 0)) %>%
  mutate(EDU.Bachelor = ifelse(EDU == "Bachelor", 1, 0)) %>%
  mutate(EDU.Graduate = ifelse(EDU == "Post-Bachelor", 1, 0)) %>%
  mutate(EDU.Other = ifelse(EDU == "Other", 1, 0)) %>%
  select(-c("X", "SERIALNO", "SPORDER", "WKHP", "TEN", "HUPAOC", "WIF", "DEAR",
            "DEYE", "DOUT", "DPHY", "DREM", "SCHL", "SEX", "ESR", "emp_type", "EDU"))

```

```{r}
if (!require(corrplot)) install.packages("corrplot")
library(corrplot)
```

```{r}
df_sub <- df_clean %>%
  select(-c("PWGTP", "NAICS"))

df_cor = cor(df_sub)
corrplot(df_cor, type="lower", tl.cex=0.7)
```

Notable correlations
NOc ~ HUPAOC.OverSix, HUPAOC.Both, HUPAOC.None
NP ~ HUPAOC.OverSix, HUPAOC.Both, HUPAOC.None
No clear correlation of absent with other variables - without oversample, only 2% were Absent

```{r}
df_oversample <- df_clean %>%
  mutate(PWGTP = ifelse(Absent == 1, ceiling(log(PWGTP)), 1)) %>%
  select(-c("NAICS"))
```

```{r}
model.base <- glm(Absent ~ ., data=as.data.frame(scale(df_sub)))
summary.base <- summary(model.base)
summary.base
```

```{r}
coefficients.base <- as.data.frame(summary.base$coefficients)
coefficients.base[-1,] %>%
  filter(`Pr(>|t|)` <= 0.05) %>%
  select(Estimate, `Pr(>|t|)`) %>%
  arrange(desc(abs(Estimate))) %>%
  head(10)

```



```{r}
oversample_weights = df_oversample$PWGTP
model.oversample <- glm(Absent ~ . - PWGTP, data = as.data.frame(scale(df_oversample)), weights = oversample_weights)
summary.oversample <- summary(model.oversample)
summary.oversample
```

```{r}
coefficients.oversample <- as.data.frame(summary.oversample$coefficients)
coefficients.oversample[-1,] %>%
  filter(`Pr(>|t|)` <= 0.05) %>%
  select(Estimate, `Pr(>|t|)`,) %>%
  arrange(desc(abs(Estimate))) %>%
  head(10)
```

